Viscous Transport in Eroding Porous Media

### 2019-07-30 docs/transport/erosion_transport.tex
working on paper

\input{preambles.tex}
\usepackage{showkeys}
\begin{document}

@@

Again, given a density function $\tau$, the boundary data of $v$ as
defined in equation~\eqref{eqn:laplaceComplex} can be found by applying
the Sokhotski-Plemelj jump relation~\eqref{eqn:SPrelation}.  The
boundary data differs if we are considering the interior problem ($x \in
\Omega$) or the exterior problem ($x \in \Omega^c$).  Therfore, for the
boundary data we use the notation $v^-$ for the interior problem and
$v^+$ for the exterior problem.  Rather than directly applying the
trapezoid rule to $v(x)$, we start with the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^{-}(y) - v(x)}{y - x} \,dy = 0,
\end{align}
where we are considering the interior problem.  Since the integrand is
bounded as $x$ approaches $\bd\Omega$, we can apply the trapezoid rule
\begin{align}
  \sum_{j=1}^{N} \frac{v^{-}(y_j) - v(x)}{y_j - x} w_j \approx 0,
\end{align}
and the error is independent of $x$.  Rearranging for $v(x)$, we have
the quadrature formula
\begin{align}
  v(x) = \frac{\sum\limits_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j}
  {\sum\limits_{j=1}^N \frac{1}{y_j - x} w_j}.
  \label{eqn:BaryvInterior}
\end{align}
Using a similar construction and letting $a$ be any point inside
$\Omega$, the Barycentric quadrature rule for the exterior problem is
\begin{align}
  v(x) = \frac{1}{x-a} 
    \frac{\sum\limits_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j}.
  \label{eqn:BaryvExterior}
\end{align}

Similar constructions can be used to derive quadrature rules
for $v'(x)$.  For $x \in \Omega$, we have the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^-(y) - v(x) + (y-x)v'(x)}{(y - x)^2} = 0,
\end{align}
and the integrand is bounded for all $x$.  Therefore, upon applying the
quadrature rule and rearranging for $v'(x)$, we have
\begin{align}
  v'(x) = \frac{\sum\limits_{j=1}^{N}
    \frac{v^{-}(y_j) - v(x)}{(y_j-x)^2} w_j}
  {\sum\limits_{j=1}^{N} \frac{1}{y_j-x} w_j}.
  \label{eqn:BaryvprimeInterior}
\end{align}
Using a similar construction, the Barycentric quadrature rule to compute
$v'(x)$ in $\Omega^c$ is
\begin{align}
  v'(x) = \frac{1}{x-a} \frac{\sum\limits_{j=1}^N
    \frac{v^+_j - v(x)}{(y_j - x)^2} w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j - x} w_j}.
  \label{eqn:BaryvprimeExterior}
\end{align}
Note that the value $v(x)$ is required to compute $v'(x)$ for both the
interior and exterior case.  These are computed using the quadrature
methods in equations~\eqref{eqn:BaryvInterior}
and~\eqref{eqn:BaryvExterior}.

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  Following the work of Ioakimidis et
al~\cite{ioa-pap-per1991} (see equation (2.12)), we mimick the
construction of the quadrature rule for $v(x)$ and $v'(x)$. We start
with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x} dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega$.  Therefore, after applying the trapezoid rule
  and solving for $v''(x)$, we have the Barycentric quadrature rule
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}.
\end{align}
Similar to the quadrature rule for $v'(x)$ requring the value $v(x)$,
the quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$.
Both these quantities are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying the Quadrature}
\label{s:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  Since the shear
stress must only be computed once per time step, and the vorticity is
only used as a post-processing step, we accept the $\mathcal{O}(N^2)$
algorithm.  However, the velocity must be computed at every GMRES
iteration, which would make the overall method impractical without a
fast algorithm.

At each GMRES iteration, we must compute the velocity at each
discretization point on $\bd\Omega$.  To reduce the cost of computing
the velocity, we start by using the trapezoid rule~\eqref{eqn:trap}
accelerated with the fast multipole method (FMM)~\cite{gre-rok1987},
which requires $\mathcal{O}(N)$ operations, and we call this velocity
$\vv_\trap(\xx)$.  For target points that are sufficiently far from
$\bd\Omega$, $\vv_\trap(\xx)$ is sufficiently accurate.  However, this
velocity is inaccurate at points that are too close to $\bd\Omega$, and
the velocity must be corrected.  Assuming that $\xx$ is too close to
$\gamma_k$, only the velocity due to $\gamma_k$ must be corrected. We
first subtract the velocity contribution of $\vv(\xx)$ due to
$\gamma_k$.  Then, the Barycentric quadrature rule is used to compute
the velocity due to $\gamma_i$ with more accuracy.  That is, the final
velocity is
\begin{align}
  \vv(\xx) = \vv_\trap(\xx) - \frac{L_k}{N_k} \sum_{j=1}^{N_\iin} 
    D(\xx,\yy^k_j) \eeta(\yy_j) + \vv^k_\bary(\xx),
\end{align}
where $\vv^k_\bary(\xx)$ is the velocity at $\xx$ resulting from
applying the Barycentric quadrature rule due to $\gamma_k$.  This method
naturally extends to points that are close to $\Gamma$, and to the
scenario when $\xx$ is simultaneously close to multiple components of
$\bd\Omega$.

Similarly, to compute the vorticity and shear stress, the trapezoid rule
is first used.  Then, if $\xx$ is close to $\gamma_k$, the trapezoid
quadrature rule contribution due to $\gamma_k$ is subtracted and the
Barycentric quadrature rule contribution due to $\gamma_k$ is added.
Since the shear stress and vorticity are only computed once per time
step, we use a direct summation rather than the FMM.  Relative to the
cost of applying GMRES with the FMM, the additional once-per-time-step
cost is minimal.

We also do not use any fast summation methods to apply the Barycentric
quadrature rule.  Since this quadrature rule is applied at every GMRES
iteration, it is imperative that we only apply the Barycentric
quadrature at target points where they are required.  As a rule of
thumb, the trapezoid rule due to $\gamma_k$ is accurate if
\begin{align}
  \label{eqn:TrapCutoff}
  d(\xx,\gamma_k) = \inf_{\yy \in \gamma_k} \|\xx - \yy\| > 
    5 \frac{L_k}{N_\iin}.
\end{align}
To avoid checking if all target point $\xx$
satisfy~\eqref{eqn:TrapCutoff}, we make use of the {\thL} variables.
First, given the two eroding bodies $\gamma_i$ and $\gamma_j$ with
lengths $L_i$ and $L_j$, and centers $\cc_i$ and $\cc_j$, we first check
if
\begin{align}
  \label{eqn:BodyCutoff}
  \|\cc_i - \cc_j\| < \frac{L_i}{2\pi} + \frac{L_j}{2\pi} + 
    \alpha_\iin \left(\frac{L_i}{N_\iin} + \frac{L_j}{N_\iin} \right),
\end{align}
where $\alpha_\iin \geq 1$ is a free parameter that needs to be chosen.
In this manner, rather than using an expensive all-to-all algorithm to
compute the distance between all pairs of discretization points, we
compute the distance between pairs of circles.  This criteria allows us
to quickly determine bodies that contain discretization points where the
Barycentric quadrature rule may need to be applied, and the parameter
$\alpha_\iin$ is accounts for the fact that the bodies are being
approximated with circles.  Assuming that the two bodies $\gamma_i$ and
$\gamma_j$ satisfy condition~\eqref{eqn:BodyCutoff}, for each point $\xx
\in \gamma_j$, we check if
\begin{align}
  \label{eqn:PointsBodyCutoff}
  \|\xx - \cc_i\| < \frac{L_i}{2\pi}
+ \alpha_\iin \left(\frac{L_i}{N_\iin} \right).
\end{align}
To determine if points on $\gamma_i$ are too close to the outer wall, we
use the fact that the outer wall is bounded in $[-3,3] \times [-1,1]$,
and that the bodies are all contained in $[-1,1] \times [-1,1]$.  We
first check if
\begin{align}
  \left\|\cc_i - \left[
    \begin{array}{c}
      x_{\cc_i } \\ \pm 1
    \end{array}
    \right]
  \right\| < \frac{L_i}{2 \pi} + \alpha_\out \frac{L_\out}{N_\out}
\end{align}
where $x_{\cc_i } $ is the x coordinate of $\cc_i$. If body $\gamma_i$
satisfies this condition, for each point $\xx=(x,y) \in \gamma_i$, we
apply the Barycentric rule to points that satisfy
\begin{align}
  \label{eqn:PointsWallCutoff}
  |y \pm 1| < \frac{L_i}{2 \pi} +\alpha_\out \frac{L_\out}{N_\out}.
\end{align}

To determine an appropriate value for $\alpha_\iin$ and $\alpha_\out$,
we first compute an accurate solution by using the Barycentric
quadrature rule for all discretization points.  Then, for multiple
values of $\alpha_\iin$ and $\alpha_\out$, we compute the velocity field
with the trapezoid rule for all points that do not satisfy
conditions~\eqref{eqn:PointsBodyCutoff}
and~\eqref{eqn:PointsWallCutoff}.  By comparing these two velocities, we
find that $\alpha_\iin = 4$ and $\alpha_\out = 4$ give sufficient
accuracy while keeping the number of points that require the expensive
Barycentric quadrature rule to a minimum.  We use these values for all
of our numerical simulations.

Finally, to determine if a target point $\xx$ that is in the fluid bulk
rather than on another eroding body requires the Barycentric quadrature
rule, we only check conditions~\eqref{eqn:PointsBodyCutoff}
and~\eqref{eqn:PointsWallCutoff}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Integration}
\label{s:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We reuse the time stepping method outlined in our previous
work~\cite{qua-moo2018}.  Rather than tracking the $(x,y)$
coordinates, the {\thL} variables are tracked. A second-order
Implicit-Explicit method is used. In particular, the diffusive term
corresponding to the curvature penalization term is discretized
implicitly, and all other terms, which are non-stiff, are treated
explicitly.

\todo[inline]{Talk about time integration for erosion and for tracers}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  For porous media, two of
the most important variables are the porosity and the permeability.
However, additional properties of the flow are revealed by computing
the anomalous diffusion rate, the tortuosity, and the distribution the
gap sizes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Diffusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dynamics of fluid in porous media can be characterized in terms of
anomalous diffusion.  However, the anomalous diffusion rate depends on
the properties of the geometry.  In addition to the porosity, the
distribution of the grains affects the anomalous diffusion rates.  To
quantify this rate, we require tracer trajectories that are governed by
\begin{align}
  \label{eqn:trajectories}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0.
\end{align}
Given a trajectory, we define $\lambda_j(t)$ to be the distance
travelled by the particle
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss_j(\tau)\|\, d\tau.
\end{align}
Having formed $N_p$ trajectories, the first and second ensemble moments
are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
If the particle spreading scales with time as $\sigma_\lambda^2 \sim
t^\alpha$, with $\alpha < 1$, this indicates that the particle
trajectories are sub-diffusive.  In contrast, if $\sigma_\lambda^2 \sim
t^2$, the particle trajectories are undergoing a ballistic motion and
are not affected by the grains.

>>>>

\input{preambles.tex}
%\usepackage{showkeys}
\begin{document}

@@

Again, given a density function $\tau$, the boundary data of $v$ as
defined in equation~\eqref{eqn:laplaceComplex} can be found by applying
the Sokhotski-Plemelj jump relation~\eqref{eqn:SPrelation}.  The
boundary data differs if we are considering the interior problem ($x \in
\Omega$) or the exterior problem ($x \in \Omega^c$).  Therfore, for the
boundary data we use the notation $v^-$ for the interior problem and
$v^+$ for the exterior problem.  Rather than directly applying the
trapezoid rule to $v(x)$, we start with the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^{-}(y) - v(x)}{y - x} \,dy = 0,
\end{align}
where we are considering the interior problem.  Since the integrand is
bounded as $x$ approaches $\bd\Omega$, we can apply the trapezoid rule
\begin{align}
  \sum_{j=1}^{N} \frac{v^{-}(y_j) - v(x)}{y_j - x} w_j \approx 0,
\end{align}
and the error is independent of $x$.  Rearranging for $v(x)$, we have
the interior Barycentric quadrature formula 
\begin{align}
  v(x) = \frac{\sum\limits_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j}
  {\sum\limits_{j=1}^N \frac{1}{y_j - x} w_j}, \quad x \in \Omega.
  \label{eqn:BaryvInterior}
\end{align}
Using a similar construction and letting $a$ be any point inside
$\Omega$, the exterior Barycentric quadrature rule is
\begin{align}
  v(x) = \frac{1}{x-a} 
    \frac{\sum\limits_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvExterior}
\end{align}

Similar constructions can be used to derive quadrature rules
for $v'(x)$.  For $x \in \Omega$, we have the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^-(y) - v(x) + (y-x)v'(x)}{(y - x)^2} = 0,
\end{align}
and the integrand is bounded for all $x \in \Omega$.  Therefore, upon
applying the quadrature rule and rearranging for $v'(x)$, we have the
interior Barycentric quadrature rule
\begin{align}
  v'(x) = \frac{\sum\limits_{j=1}^{N}
    \frac{v^{-}(y_j) - v(x)}{(y_j-x)^2} w_j}
  {\sum\limits_{j=1}^{N} \frac{1}{y_j-x} w_j}, 
  \quad x \in \Omega.
  \label{eqn:BaryvprimeInterior}
\end{align}
Using a similar construction, the exterior Barycentric quadrature rule
for the the first derivative is
\begin{align}
  v'(x) = \frac{1}{x-a} \frac{\sum\limits_{j=1}^N
    \frac{v^+_j - v(x)}{(y_j - x)^2} w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j - x} w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvprimeExterior}
\end{align}
Note that the value $v(x)$ is required to compute $v'(x)$ for both the
interior and exterior case.  These are computed using the quadrature
methods in equations~\eqref{eqn:BaryvInterior}
and~\eqref{eqn:BaryvExterior}.

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  Following the work of Ioakimidis et
al~\cite{ioa-pap-per1991} (see equation (2.12)), we mimick the
construction of the quadrature rule for $v(x)$ and $v'(x)$. We start
with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x} dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivate
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
Similar to the quadrature rule for $v'(x)$ requring the value $v(x)$,
the quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$.
Both these quantities are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying the Quadrature}
\label{s:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  By using a fast
algorithm such as the fast multipole method~\cite{gre-rok1987}, this
cost can be reduced to $\mathcal{O}(N)$ operations.  However, the
Barycentric quadrature formulas require several $N$-body calculations
which we would like to avoid.  Instead, we use a hybrid method where the
fast multipole accelerate trapezoid rule is used to compute the velocity
at most of the discretization points, and the non-accelerated
Barycentric quadrature rule is used to locally correct the trapezoid
rule.

We start by using the trapezoid rule~\eqref{eqn:trap} accelerated with
the FMM, which requires $\mathcal{O}(N)$ operations, and we call this
velocity $\vv_\trap(\xx)$.  For target points that are sufficiently far
from $\bd\Omega$, $\vv_\trap(\xx)$ is sufficiently accurate.  However,
if $\xx$ is sufficiently close to one of the eroding bodies $\gamma_k$,
or the outer wall $\Gamma$, the velocity contribution from that
particular component of $\bd\Omega$ is inaccurate and it must be
corrected.  Assuming that $\xx$ is too close to $\gamma_k$, we first
subtract the velocity contribution of $\vv(\xx)$ due to $\gamma_k$.
Then, the Barycentric quadrature rule is used to compute the velocity
due to $\gamma_k$ with more accuracy.  That is, the final velocity is
\begin{align}
  \vv(\xx) = \vv_\trap(\xx) - \frac{L_k}{N_k} \sum_{j=1}^{N_\iin} 
    D(\xx,\yy^k_j) \eeta(\yy_j) + \vv^k_\bary(\xx),
  \label{eqn:velocityDecomp}
\end{align}
where $\vv^k_\bary(\xx)$ is the velocity at $\xx$ resulting from
applying the Barycentric quadrature rule due to $\gamma_k$.  This
strategy naturally extends to points that are close to $\Gamma$, and to
the scenario when $\xx$ is simultaneously close to multiple components
of $\bd\Omega$.  While the term $\vv_\trap(\xx)$ in
equation~\eqref{eqn:velocityDecomp} is computed for all target points
using the FMM, the other two terms are computed with a direct summation.
However, these terms are only required for target points near an eroding
body or the outer wall, and the number of such points is much less than
$N$.  An identical quadrature method is used to compute the vorticity
and shear stress.  That is, these quantities are computed using the
trapezoid rule, and then updates due to inaccuracies of the trapezoid
rule are made.  However, since the shear stress and vorticity are only
computed once per time step, the trapezoid rule is applied with a direct
summation rather than the FMM.  Relative to the cost of applying GMRES
with the FMM, the additional once-per-time-step cost is minimal.

Per grid point, evaluating the Barycentric quadrature rules dominates
the computational cost. Therefore, it is imperative that we only apply
the Barycentric quadrature when the trapezoid rule is inaccurate.  As a
rule of thumb, the trapezoid rule due to $\gamma_k$ is
accurate~\cite{bar2014} if
\begin{align}
  \label{eqn:TrapCutoff}
  d(\xx,\gamma_k) = \inf_{\yy \in \gamma_k} \|\xx - \yy\| > 
    5 \frac{L_k}{N_\iin}.
\end{align}
To avoid checking if all target point $\xx$
satisfy~\eqref{eqn:TrapCutoff}, we make use of the {\thL} variables.
First, given the two eroding bodies $\gamma_i$ and $\gamma_j$ with
lengths $L_i$ and $L_j$, and centers $\cc_i$ and $\cc_j$, we first check
if
\begin{align}
  \label{eqn:BodyCutoff}
  \|\cc_i - \cc_j\| < \frac{L_i}{2\pi} + \frac{L_j}{2\pi} + 
    \alpha_\iin \left(\frac{L_i}{N_\iin} + \frac{L_j}{N_\iin} \right),
\end{align}
where $\alpha_\iin \geq 1$ is a parameter that needs to be chosen.  In
this manner, rather than using an expensive all-to-all algorithm to
compute the distance between all pairs of discretization points, we
compute the distance between pairs of circles.  This criteria allows us
to quickly determine bodies that contain discretization points where the
Barycentric quadrature rule may need to be applied, and the parameter
$\alpha_\iin$ accounts for the fact that the bodies are being
approximated with circles.  Assuming that the two bodies $\gamma_i$ and
$\gamma_j$ satisfy condition~\eqref{eqn:BodyCutoff}, for each point $\xx
\in \gamma_j$, we check if
\begin{align}
  \label{eqn:PointsBodyCutoff}
  \|\xx - \cc_i\| < \frac{L_i}{2\pi}
+ \alpha_\iin \left(\frac{L_i}{N_\iin} \right).
\end{align}
To determine if points on $\gamma_i$ are too close to the outer wall, we
use the fact that the outer wall is bounded in $[-3,3] \times [-1,1]$,
and that the bodies are all contained in $[-1,1] \times [-1,1]$.  We
first check if
\begin{align}
  \left\|\cc_i - \left[
    \begin{array}{c}
      x_i \\ \pm 1
    \end{array}
    \right]
  \right\| < \frac{L_i}{2 \pi} + \alpha_\out \frac{L_\out}{N_\out}
\end{align}
where $x_i$ is the x coordinate of $\cc_i$. If body $\gamma_i$ satisfies
this condition, for each point $\xx=(x,y) \in \gamma_i$, we apply the
Barycentric rule to points that satisfy
\begin{align}
  \label{eqn:PointsWallCutoff}
  |y \pm 1| < \frac{L_i}{2 \pi} +\alpha_\out \frac{L_\out}{N_\out}.
\end{align}
Finally, to determine if a target point $\xx$ that is in the fluid bulk
rather than on another eroding body requires the Barycentric quadrature
rule, we only check conditions~\eqref{eqn:PointsBodyCutoff}
and~\eqref{eqn:PointsWallCutoff}.

To determine an appropriate value for $\alpha_\iin$ and $\alpha_\out$,
we fix a eroded geometry and we compute an accurate solution by using
the Barycentric quadrature rule for all discretization points.  Then,
for multiple values of $\alpha_\iin$ and $\alpha_\out$, we compute the
velocity field with the trapezoid rule for all points that do not
satisfy conditions~\eqref{eqn:PointsBodyCutoff}
and~\eqref{eqn:PointsWallCutoff}.  By comparing these two velocities, we
find that $\alpha_\iin = 4$ and $\alpha_\out = 4$ give sufficient
accuracy while keeping the number of points that require the expensive
Barycentric quadrature rule to a minimum.  We use these values for all
of our numerical simulations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Integration}
\label{s:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We reuse the time stepping method outlined in our previous
work~\cite{qua-moo2018}.  Rather than tracking the $(x,y)$ coordinates,
the {\thL} variables are tracked. In addition, a tangential velocity
field is added to each body so that the discretization points remain
equispaced for all time.  A second-order Implicit-Explicit method is
used. In particular, the diffusive term corresponding to the curvature
penalization term is discretized implicitly, and all other terms, which
are non-stiff, are treated explicitly.  By using this time stepping
method in conjunction with the curvature penalization and Gaussian
filter, we stably evolve the eroding bodies.   An in depth description
of the time stepping method for erosion is in Section 3.3 of our
previous work~\cite{qua-moo2018}.

In Section~\ref{s:transport}, we we analyze passive tracers suspended in
an eroded geometry.  Give such a geometry $\Omega$, a tracer $\ss(t)$
initialized at $\ss_0$ is governed by 
\begin{align}
  \label{eqn:trajectories}
  \frac{d\ss}{dt} = \uu(\ss), \quad \ss(0) = \ss_0.
\end{align}
To compute such a trajectory, we apply the standard fourth-order
Runge-Kutta method.  To evaluate $\uu(\ss)$, depending on the proximity
of $\ss$ to $\bd\Omega$, we either use the trapezoid rule or the
Barycentric quadrature rule as described in Section~\ref{s:fmm}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{s:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  For porous media, two of
the most important variables are the porosity and the permeability.
However, additional properties of the flow are revealed by computing
the anomalous diffusion rate, the tortuosity, and the distribution the
gap sizes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Diffusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dynamics of fluid in porous media can be characterized in terms of
anomalous diffusion.  However, the anomalous diffusion rate depends on
the properties of the geometry.  In addition to the porosity, the
distribution of the grains affects the anomalous diffusion rates.  To
quantify this rate, we require tracer trajectories that are governed by
equation~\eqref{eqn:trajectories}.  Given a trajectory, we define
$\lambda_j(t)$ to be the distance travelled by the particle
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Having formed $N_p$ trajectories, the first and second ensemble moments
are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
If the particle spreading scales with time as $\sigma_\lambda^2 \sim
t^\alpha$, with $\alpha < 1$, this indicates that the particle
trajectories are sub-diffusive.  In contrast, if $\sigma_\lambda^2 \sim
t^2$, the particle trajectories are undergoing a ballistic motion and
are not affected by the grains.
